---
title: "Solutions to Hadley Wickham's ggplot2: Elegant graphics for Data Analysis 2nd ed."
author: "Alfred Johnson"
Date:
output:
  html_document:
    toc: true
    toc_float: true
---

#2.2.1 Exercises

```{r}
library(tidyverse)
```
1. List five functions that you could use to get more information about the
mpg dataset.
```{r}
#description of the dataset.
str(mpg)
```
```{r}
#check the first n rows of data from the dataset.
head(mpg,6)
```

```{r}
#check last n rows of the dataset.
tail(mpg,6)
```

```{r}
#check the dimensions of the dataframe.
dim(mpg)
```

```{r}
#get summary stats from the dataset.
summary(mpg)
```

2.How can you find out what other datasets are included with ggplot2?
```{r}
#syntax holds true to get datasets available in any package.
data(package = "ggplot2")
```

3.Apart from the US, most countries use fuel consumption (fuel consumed
over fixed distance) rather than fuel economy (distance travelled with fixed
amount of fuel). How could you convert cty and hwy into the European
standard of l/100km?

```{r}
# 1 mpg = 235.21 l/100km (https://www.unitjuggler.com/convert-fuelconsumption-from-mpg-to-lper100km.html)
mpg%>%mutate(cty = cty*235.21,hwy = hwy*235.21)
```
4. Which manufacturer has the most the models in this dataset? Which model
has the most variations? Does your answer change if you remove the redundant specification of drive train (e.g. “pathfinder 4wd”, “a4 quattro”)
from the model name?
```{r}
#manufactorer with most models
mpg%>%count(manufacturer,sort = T)
```


```{r}
#most models in dataset. Most variations would mean the distinct count of number of times a model is repeated.
mpg%>%count(model,sort = T)
```
```{r}
#remove the reduntant specs
mpg%>%mutate(model=str_replace_all(model,c("quattro" = "","2wd" = "","4wd" = "")))
```
```{r}
#Rechecking the max variation in models.
mpg%>%
  mutate(model=str_replace_all(model,c("quattro" = "","2wd" = "","4wd" = "")))%>%
  mutate_if(is.character,trimws)%>%
  count(model,sort = T)

#it is important to trim when using str_replace as it does not deal with any whitespaces left and count will condsider "a4" and "a4 " as two distinct values.

```

#2.3.1 Exercises
1. How would you describe the relationship between cty and hwy? Do you
have any concerns about drawing conclusions from that plot?
```{r}
#plot a scatter between cty and hwy.
ggplot(mpg,aes(cty,hwy)) +
  geom_point()
```
The relationship looks like a linear one. With increaseing values for cty we get greater values for hwy as well.

2. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
Is it useful? How could you modify the data to make it more informative?

```{r}
ggplot(mpg,aes(model,manufacturer)) +
  geom_point()
```

it would be better if we could rotate the x axis able to make them visible and not overlap.


3. Describe the data, aesthetic mappings and layers used for each of the
  following plots. You’ll need to guess a little because you haven’t seen all
  the datasets and functions yet, but use your common sense! See if you can
  predict what the plot will look like before running the code.
  1: ggplot(mpg, aes(cty, hwy)) + geom_point()
  2: ggplot(diamonds, aes(carat, price)) + geom_point()
  3: ggplot(economics, aes(date, unemploy)) + geom_line()
  4: ggplot(mpg, aes(cty)) + geom_histogram()
  
```{r}
#scatter plot for x=cty and y=hwy
ggplot(mpg,aes(cty,hwy)) +
  geom_point()
```
  
```{r}
#scatter plot using the diamond dataset where x=carat,y=price
ggplot(diamonds,aes(carat,price)) +
 geom_point()
```
```{r}
#line plot using the economic dataset.
ggplot(economics,aes(date,unemploy)) +
  geom_line()
```
```{r}
#histogram made from the mpg dataset on the cty variable.
ggplot(mpg,aes(cty)) +
  geom_histogram()
```

#2.4.1 Exercises

1. Experiment with the colour, shape and size aesthetics. What happens
when you map them to continuous values? What about categorical values?
What happens when you use more than one aesthetic in a plot?

```{r}
summary(mpg)
```

we can see variables like cty,hwy,year are some continuous variables and model,trans,drv,class are categorical.
```{r}
#using a continuous variable you can see that the scale becomes non-discrete.Adding multiple values in aes will give us more than one aesthetic liek color as well as size in the below plot.
ggplot(mpg,aes(displ,hwy,color = year,size = cyl)) +
  geom_point()
```

2. What happens if you map a continuous variable to shape? Why? What
happens if you map trans to shape? Why?
```{r}
ggplot(mpg,aes(displ,hwy,shape = trans)) +
  geom_point()
```
The shape aesthetic cannot accept more than 6 discrete values. Any more than 6 are not displayed and the points are removed from the scatter plot. No continuous variable is accepted by the shape parameter.

3.How is drive train related to fuel economy? How is drive train related to
engine size and class?
```{r}
#drive train with engine displacement and added color using class.
ggplot(mpg,aes(drv,displ,color = class)) +
  geom_point()
```
```{r}
#drive train with fuel economy
ggplot(mpg,aes(drv,hwy,color = class)) +
  geom_point()
```

#2.5.1


